---
title: "SurvivalHW1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, we will install/load packages and read in the data.

```{r}
install.packages("survival")
install.packages("survminer")
library(dplyr)
library(survival)
library(survminer)
library(haven)
hurricane = read_sas("C:\\Users\\sydne\\Documents\\Survival Analysis\\Homework1_SA\\hurricane.sas7bdat")
```

next we will find the percentage of pumps that survived the hurricane overall


```{r}
mean(hurricane$survive)
```

then we will group by the reason of failure to find the percent that survived, the percentage of observations and the average failure time by reason of failure.


```{r}

hurricane %>%
  group_by(reason) %>%
  summarise(percent_surv = mean(survive), percent = n()/770, avg_failtime = mean(hour))


```


ANOVA and equal vairances test

```{r}
library(lawstat)
levene.test(hurricane$hour, hurricane$reason)

pre_anova<-lm(hour~factor(reason), data = hurricane)
summary(pre_anova)
anova(pre_anova)

```


Survival probability plot (not stratified)


```{r}
recid_surv <- Surv(time = hurricane$hour, event = hurricane$survive == 0)

recid_km <- survfit(recid_surv ~ 1, data = hurricane)
summary(recid_km)
plot(recid_km, main = "Survival Function", xlab = "hour", ylab = "Survival Probability")

ggsurvplot(recid_km, data = hurricane, conf.int = TRUE, palette = "purple",
           xlab = "hour", ylab = "Survival Probability", legend = "none",
           break.y.by = 0.1)
```

stratified survival probability plot

```{r}


survdiff(recid_surv ~ reason, rho = 0, data = hurricane)

recid_strat <- survfit(recid_surv ~ reason, data = hurricane)
ggsurvplot(recid_strat, data = hurricane, pval = TRUE)

```


hazard function plot and calculations

```{r}
recid_km$hp <- recid_km$n.event/recid_km$n.risk
recid_haz <- merge(data.frame(time = seq(1,48,1)), data.frame(time = recid_km$time, hp = recid_km$hp), by = "time", all = TRUE)
recid_haz[is.na(recid_haz) == TRUE] <- 0

plot(y = recid_haz$hp, x = recid_haz$time, main = "Hazard Probability Function", xlab = "Tenure", ylab = "Hazard Probability",
     type = 'l')

ggsurvplot(recid_km, data = hurricane, fun = "cumhaz", conf.int = TRUE, palette = "purple",
           xlab = "Hour", ylab = "Cumulative Hazard", legend = "none")

```

















