---
title: "R Notebook"
output: html_notebook
---

```{r}
library(haven)
library(forecast)
library(fma)
library(tseries)
library(expsmooth)
library(lmtest)
library(zoo)
library(ggplot2)
library(caschrono)
library(TSA)
library(quantmod)
library(dplyr)
library(MASS)
```



```{r}
ts_data<-read.csv("C:\\Users\\sydne\\Downloads\\TS_proj_agg.csv")

ts_trainData<-ts(ts_data[1:54,'PM25'], start = 2014, frequency= 12)
ts_testData<-ts(ts_data[55:60,'PM25'], start = c(2018,7), frequency = 12)
```


```{r}
arima.1<-Arima(ts_trainData,order=c(0,0,0), seasonal= c(0,0,0),xreg=fourier(ts_trainData,K=4))
summary(arima.1)
plot.ts(arima.1$residuals)
Acf(arima.1$residuals)
arima.1$residuals %>%
  ggtsdisplay()
```

ESM FORECASTS

```{r}
fc <- holt(ts_trainData,initial = "optimal", h=6)
fc2 <- holt(ts_trainData, damped=TRUE, h=6)
fc3 <- hw(ts_trainData, seasonal = "additive", h=6)
fc4 <- hw(ts_trainData, seasonal = "multiplicative", h=6)


plot(fc, main = "Daily concentrations -Simple ESM", xlab = "Date", ylab = "Average Daily Concentrations")
abline(v = 2014, col = "red", lty = "dashed")


plot(fc2, main = "Daily concentrations-Damped ESM", xlab = "Date", ylab = "Average Daily Concentrations")
abline(v = 2014, col = "red", lty = "dashed")

plot(fc3, main = "Daily concentrations-Additive Seasonality", xlab = "Date", ylab = "Average Daily Concentrations")
abline(v = 2014, col = "red", lty = "dashed")

plot(fc4, main = "Daily concentrations-Multiplicative Seasonality", xlab = "Date", ylab = "Average Daily Concentrations")
abline(v = 2014, col = "red", lty = "dashed")
```

```{r}
test.results=forecast(fc,h=6)
test.results$mean
test.results
error=ts_testData-test.results$mean
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE

test.results=forecast(fc2,h=6)
test.results$mean
test.results
error=ts_testData-test.results$mean
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE

test.results=forecast(fc3,h=6)
test.results$mean
test.results
error=ts_testData-test.results$mean
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE

#THE FORECAST BELOW (FC4) ENDED UP BEING CHOSEN

test.results=forecast(fc4,h=6)
test.results$mean
test.results
error=ts_testData-test.results$mean
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE
```

ARIMA WITH NO SEASON MODEL

```{r}
auto.arima(ts_trainData)
Arima.1<-Arima(ts_trainData, order = c(2,1,0) )
Acf(Arima.1$residuals)
White.LB <- rep(NA, 10)
for(i in 1:50){
  White.LB[i] <- Box.test(Arima.1$residuals, lag = i, type = "Ljung", fitdf = 2)$p.value
}

White.LB <- pmin(White.LB, 0.2)
barplot(White.LB, main = "Ljung-Box Test P-values", ylab = "Probabilities", xlab = "Lags", ylim = c(0, 0.2))
abline(h = 0.01, lty = "dashed", col = "black")
abline(h = 0.05, lty = "dashed", col = "black")
```


FORECAST AND MAPE FOR ARIMA WITH NO SEASON


```{r}
test.results1=forecast(Arima.1, h=6)
results = ts(data.frame(test.results1)$Point.Forecast, start = c(2018,7), frequency = 12)
ts_trainData
error=ts_testData-results
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE
```

ARIMA WITH SEASON

```{r}
Arima.2<-Arima(ts_trainData, order = c(2,0,0), xreg = fourier(ts_trainData, K=4))
Acf(Arima.2$residuals)
White.LB <- rep(NA, 10)
for(i in 1:50){
  White.LB[i] <- Box.test(Arima.1$residuals, lag = i, type = "Ljung", fitdf = 2)$p.value
}

White.LB <- pmin(White.LB, 0.2)
barplot(White.LB, main = "Ljung-Box Test P-values", ylab = "Probabilities", xlab = "Lags", ylim = c(0, 0.2))
abline(h = 0.01, lty = "dashed", col = "black")
abline(h = 0.05, lty = "dashed", col = "black")
```

FORECAST AND MAPE FOR ARIMA WITH SEASON - THIS WAS THE WINNING MODEL

```{r}
test.results2=forecast(Arima.2, xreg = fourier(ts_trainData, K=4, h=6))
results2 = data.frame(test.results2)$Point.Forecast
ts_testData
results2=ts(results2, start = c(2018,7), frequency = 12)
error=ts_testData-results2
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE
```


BELOW WE MADE ALL OF THE PROVIDED DATA OUR TRAINING SET AND THEN PREDICTED 2019

```{r}
new_train<-ts(ts_data$PM25, start= 2014, frequency = 12 )
full_arima <- Arima(new_train, order = c(2,0,0), xreg = fourier(new_train, K=4))
forecast2019 <- forecast(full_arima, xreg = fourier(new_train, K=4, h=6))
forecast2019$mean
```


BELOW IS THE ENSEMBLE MODEL THAT AVERAGED UCM, ESM, ARIMA WITH SEASON, AND ARIMA WITHOUT SEASON


```{r}
ensemble<-as.data.frame(cbind(c(9.232852, 9.254682, 8.712030, 9.268077, 8.865813, 9.619317), results, test.results$mean, results2))
colnames(ensemble)<-c("ARIMAX", "ARIMA (no seas)", "MultSeasESM", "ARIMA (seas)")

ensemble<-ensemble %>%
  mutate(ensemble = (ARIMAX + `ARIMA (no seas)` + MultSeasESM + `ARIMA (seas)`)/4)

ensMod<-ts(ensemble$ensemble, start=c(2018,7), frequency = 12)
error=ts_testData-ensMod
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE

```

BELOW IS TRIAL AND ERROR FOR NEURAL NET


```{r}

ADF.Pvalues <- rep(NA, 3)
for(i in 0:2){
  ADF.Pvalues[i+1] <- adf.test(ts_trainData, alternative = "stationary", k = i)$p.value
}

ADF.Pvalues


# Autoregressive Neural Network Model #
NN.Model <- nnetar(diff(ts_trainData,1), p = 2, size = 3)
NN.Forecast <- forecast(NN.Model, h = 6)
plot(NN.Forecast)

xreg1<-cbind(fourier(ts_trainData,K=4),seq(1,length(ts_trainData)),(seq(1,length(ts_trainData)))^2)
colnames(xreg1)<-c('s1','c1','s2','c2','s3','c3','s4','c4','time', 'timesq')
Model.four<-Arima(ts_trainData,order=c(0,0,0),xreg=xreg1)
NN.Model2<-nnetar(Model.four$residuals,p=2,P=1,size=2)
NN.Forecast2<-forecast(NN.Model2,h=6)
plot(NN.Forecast2)


for.seq<-seq(55,60)
xreg2<-cbind(fourier(ts_trainData,K=4,h=6),for.seq, for.seq^2)
colnames(xreg2)<-c('s1','c1','s2','c2','s3','c3','s4','c4','time', 'timesq')
Base.forecast<-forecast(Model.four,xreg=xreg2,h=6)
net.Forecast2 <- Base.forecast$mean+NN.Forecast2$mean

net.Forecast2 <- ts(net.Forecast2, start = c(2018, 7), frequency = 12)

plot(ts_testData, main = "US Airline Passengers ARIMA Model Forecasts", xlab = "Date", ylab = "Passengers (Thousands)", xlim = c(2018, 2019), ylim = c(0,20))
lines(net.Forecast2, col = "orange")
abline(v = 2008.25, col = "red", lty = "dashed")
```


```{r}
error=ts_testData-net.Forecast2
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE
```

STILL TRIAL AND ERROR FOR NEURAL NET


```{r}
set.seed(123)
MAPE<-rep(NA,40)
for(i in 1:40){
  xreg1<-cbind(fourier(ts_trainData,K=4),c(1:54),c(1:54)^2)
colnames(xreg1)<-c('s1','c1','s2','c2','s3','c3','s4','c4','time', 'timesq')
Model.four<-Arima(ts_trainData,order=c(0,0,0),xreg=xreg1)
NN.Model2<-nnetar(Model.four$residuals,p=11,P=1,size=i)
NN.Forecast2<-forecast(NN.Model2,h=6)
plot(NN.Forecast2)

xreg2<-cbind(fourier(ts_trainData,K=4,h=6),c(55:60), c(55:60)^2)
colnames(xreg2)<-c('s1','c1','s2','c2','s3','c3','s4','c4','time', 'timesq')
Base.forecast<-forecast(Model.four,xreg=xreg2,h=6)
net.Forecast2 <- Base.forecast$mean+NN.Forecast2$mean

net.Forecast2 <- ts(net.Forecast2, start = c(2018, 7), frequency = 12)
error=ts_testData-net.Forecast2
MAE=mean(abs(error))
MAPE[i]=mean(abs(error)/abs(ts_testData))
}

plot(MAPE, label = TRUE)
```

FORECAST AND MAPE FOR NEURAL NET

```{r}
set.seed(123)
xreg1<-cbind(fourier(ts_trainData,K=4),c(1:54),c(1:54)^2)
colnames(xreg1)<-c('s1','c1','s2','c2','s3','c3','s4','c4','time', 'timesq')
Model.four<-Arima(ts_trainData,order=c(0,0,0),xreg=xreg1)
NN.Model2<-nnetar(Model.four$residuals,p=11,P=1,size=3)
NN.Forecast2<-forecast(NN.Model2,h=6)
plot(NN.Forecast2)

xreg2<-cbind(fourier(ts_trainData,K=4,h=6),c(55:60), c(55:60)^2)
colnames(xreg2)<-c('s1','c1','s2','c2','s3','c3','s4','c4','time', 'timesq')
Base.forecast<-forecast(Model.four,xreg=xreg2,h=6)
net.Forecast2 <- Base.forecast$mean+NN.Forecast2$mean

net.Forecast2 <- ts(net.Forecast2, start = c(2018, 7), frequency = 12)
error=ts_testData-net.Forecast2
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))

MAPE
```


BELOW WE ADDED IN THE NEURAL NET TO THE ENSEMBLE MODEL 



```{r}
ensemble<-as.data.frame(cbind(c(9.232852, 9.254682, 8.712030, 9.268077, 8.865813, 9.619317), results, results2, net.Forecast2))
colnames(ensemble)<-c("ARIMAX", "ARIMA (no seas)", "ARIMA (seas)", "NN3")

ensemble<-ensemble %>%
  mutate(ensemble = (ARIMAX + `ARIMA (no seas)` + `ARIMA (seas)` + net.Forecast2)/4)

ensMod<-ts(ensemble$ensemble, start=c(2018,7), frequency = 12)
error=ts_testData-ensMod
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE
```

THIS NEXT PART WAS JACKSONS CODE FOR ARIMAX I JUST ADDED IN THE FOURIER TERMS

```{r}
df<-read.csv("C:\\Users\\sydne\\Downloads\\TS_proj_agg.csv")
Y_t <- ts(df$PM25[1:54],frequency=12)
X_ts <- as.matrix(df[1:54,4:14])
Y_ttest <- ts(df$PM25[55:60])
 
########################
#Plot and look at ACF, PACF
########################
plot(Y_t)
acf(Y_t, lag.max=24)
pacf(Y_t, lag.max=24)

########################
#Basic ARIMA
########################
model.1 <- auto.arima(Y_t)
summary(model.1)
acf(model.1$residuals, lag.max=24)
pacf(model.1$residuals, lag.max=24)
Box.test(model.1$residuals, lag = 24, type = c("Ljung-Box"))

########################
#Find best predictors
########################
# Fit the full model 
full.model <- lm(df$PM25 ~ df$AWND + df$SNOW + df$SNWD + df$TAVG + df$TMIN + df$TMAX
                 + df$WSF2 + df$WSF5 + df$CO + df$SO + df$NO, data = df)
# Stepwise regression model
step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
summary(step.model)
#looks like CO, SO and TAVG are the best


#HERE IS WHERE I ADDED IN SEASON
X_ts <- as.matrix(cbind(fourier(ts_trainData, K=4), df[1:54,c(7,12,13)]))
model.2 <- auto.arima(Y_t, xreg=X_ts)
summary(model.2)
Acf(model.2$residuals, lag.max=24)
Pacf(model.2$residuals, lag.max=24)
Box.test(model.2$residuals, lag = 24, type = c("Ljung-Box"))
White.LB <- rep(NA, 10)
for(i in 1:24){
  White.LB[i] <- Box.test(model.2$residuals, lag = i, type = "Ljung", fitdf = 1)$p.value
}
White.LB <- pmin(White.LB, 0.2)
barplot(White.LB, main = "Ljung-Box Test P-values", ylab = "Probabilities", xlab = "Lags", ylim = c(0, 0.2))
abline(h = 0.01, lty = "dashed", col = "black")
abline(h = 0.05, lty = "dashed", col = "black")

```
FORECAST AND MAPE FOR ARIMAX

```{r}
X_ts <- as.matrix(cbind(fourier(ts_trainData, K=4, h=6), df[55:60,c(7,12,13)]))
arimax.fc<-forecast(model.2, xreg = X_ts, h =6 )
arimax.ts<-ts(arimax.fc$mean, start = c(2018, 7), frequency = 12)
error=ts_testData-arimax.ts
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE
arimax.ts
```

THE PREDICTIONS JACKSON GAVE ME FROM UCM 

```{r}
ucm.preds<-ts(c(10.04430144,
10.10967248,
6.955645137,
9.528910279,
9.868727247,
11.16931372), start = c(2018,7), frequency = 12)
```


ADDING IN UCM TO ENSEMBLE MODEL BUT STILL NOT AS GOOD

```{r}
ensemble<-as.data.frame(cbind(arimax.ts, results, results2, net.Forecast2, test.results$mean, ucm.preds))
colnames(ensemble)<-c("ARIMAX", "ARIMA (no seas)", "ARIMA (seas)", "NN3", "ESM","UCM" )

ensemble<-ensemble %>%
  mutate(ensemble = (ARIMAX + `ARIMA (no seas)` + `ARIMA (seas)` + net.Forecast2+ESM+UCM)/6)

ensMod<-ts(ensemble$ensemble, start=c(2018,7), frequency = 12)
error=ts_testData-ensMod
MAE=mean(abs(error))
MAPE=mean(abs(error)/abs(ts_testData))
MAE
MAPE
```







